---
title: "lab 2"
author: "Eric Börjesson, Ludvig Ekman, Wenli Zhang, Tim Grube"
date: "17/11/2020"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
library(tidyverse) 
library(car) 
library(knitr)
library(psych)
library(pwr)
library(stargazer)
```

## Time
### Question a.

*Long text*


```{r 1a-vector}
data <- read.csv("./gotaflix-abn.csv")

pwr::pwr.anova.test(k=5, n=80, f = 0.0621, sig.level = 0.05)

```

### Question b.

```{r 1b}
data <- read.csv("./gotaflix-abn.csv")

data$Cover <- as.factor(data$Cover)
data$Engagement <- as.numeric(data$Engagement) 

aggregate(Engagement~Cover, data, mean) #Cover(grouping),calculating the value of Engagement in one group
aggregate(Engagement~Cover, data, median)
aggregate(Engagement~Cover, data, var)
aggregate(Engagement~Cover, data, sd) 

psych::describeBy(data, group = data$Cover) 
```

### Question c.

```{r c}
lindata <- lm(Engagement ~ Cover, data)
```
The intercept is the average engagement for the original cover. 

What is the value the model gives when only the coefficient for Cover C equals to 1
and the rest of the coefficients equals to 0?
Intercept + CoverC 
0.16 + 0.017..


### Question d.
What needs to be normal? 
Residuals

Does the S-W test agree with the qqplot? 
Both the qq-plot and the s-w test indicates that the residuals are normal. 
```{r d}

qqPlot(lindata)
plot(lindata)
shapiro.test(data$Engagement) 

```

### Question e.

If the assumption of homoscedasticity would not be fulfilled, the scatter plot would show that the variance is higher for some x values. For example, higher x values could lead to a higher variance of residuals. Looking at the following scatter plot, this is not the case. Instead, the variance is very similar and there just very few outliers.

```{r e1, fig.keep='last'}
plot(lindata)
```

The Levene test shows, that the F value is a lot lower than the critical F value 2.37 and the p-value is a lot higher than 0.05. Therefore, the null hypothesis cannot be rejected, so there is still no evidence that it is not homoscedastic.

```{r e2}

leveneTest(lindata)

```

### Question f.

The independence of the residuals is based on the design of the experiment and should be verified here. In this case, the users were randomly allocated into experimental groups and no one is asked twice, so the data should be independent. However, it can be checked additionally with a Durbin-Watson-test for example. The p-value is a lot higher than 0.05, so the null hypothesis cannot be rejected and the test indicates as well, that the residuals are independent.

```{r f}
durbinWatsonTest(lindata)
```

### Question g.

Import of data and calculating of linear model:

```{r g1}
dataMod <- read.csv("./gotaflix-abn-modified.csv")

dataMod$Cover <- as.factor(dataMod$Cover)
dataMod$Engagement <- as.numeric(dataMod$Engagement) 

lindataMod <- lm(Engagement ~ Cover, dataMod)
```

Looking at the following scatter plot, this time there are huge differences between the variances of the different X values, especially for cover C, but also cover E. This is already a strong indication that this data is not homoscedastic.

```{r g2, fig.keep='last'}
plot(lindataMod)
```

The levene test supports this statement. The F value is a lot over the critical F value and the p-value much smaller than 0.05. Therfore, the null hypothesis has to be rejected and the data is not homoscedastic.

```{r g3}
leveneTest(lindataMod)
```

### Question h. 

The ANOVA table shows that the F value is 5.4844 which is higher than the critical F value of 2.37, so the model is statistically significant. For calculating the F value, the mean squares for the covers (0.0588) and the mean squares for the residuals (0.0107) are determined first and then are divided.

``` {r h1}
Anova(lindata)
```

The results of the Tukey HSD test (HSD = honestly significant difference) show that there are three pairs of covers which have a significant difference. Cover C is the only cover which improves the engagement, at least in comparison to the covers A, B and D. There is no significant difference to cover E.

``` {r h2}
TukeyHSD(aov(lindata))
```



## Full factorial experiment
### Question a.
*How many and what are the experimental groups?*  
There are 2 different factors and each factor has 2 levels, so there are total of 4 experimental groups. The groups are:

1. Cover = Character & Summary = Character
2. Cover = Character & Summary = Genre
3. Cover = Genre & Summary = Character
4. Cover = Genre & Summary = Genre

### Question b.
*The design of this experiment can be seen as a two-way ANOVA. Load the data (gotaflix-2wayANOVA.csv) and fit a linear model. Write the equation of this linear model. What does it mean the intercept value?*  
The intercept is the models estimation of the engagement value when the two factors Cover and Summary are set to 0. In this case, that is when *Cover = Character & Summary = Character*. 
We have created 3 linear models below: lm.1, lm.2 and lm.3. The model lm.1 uses the two independent variables *Cover & Summary* to predict the dependent variable Engagement. The models lm.2 and lm.3 also includes the interaction between the factors and are modeled in two different ways. An inspection of the results below show that they produce equivalent outcome. 
```{r twoWayAnova}
twoWayAnova <- read.csv("./gotaflix-2wayANOVA.csv", stringsAsFactors = TRUE)
lm1 <- lm(Engagement ~ Cover + Summary, data = twoWayAnova)
lm2 <- lm(Engagement ~ Cover + Summary + Cover:Summary, data = twoWayAnova)
lm3 <- lm(Engagement ~ Cover*Summary, data = twoWayAnova)
summary(lm1) # alpha = 0.169028
summary(lm2) # alpha = 0.175266
summary(lm3) # alpha = 0.175266
```
### Question c.
*Analyze the assumptions of this two-way ANOVA. Independence, normality and homoscedasticity*  
**Independence** - The assumption of independence means that there should not exist any relationship between the observations within and between the groups. This is related to the design of the study and is not something we can verify from the given data set. However from the provided description, 50 movies have been randomly selected and users have been randomly assigned to the experimental groups. This would conform with the assumption of independence since both the movies and users are randomly picked. However, the movies have been randomly selected among the top 200 movies and is thereby not a random sample of the population. This could effectively influence the outcome and the result might not be possible to generalize for the entire population, that is all the movies.  

**Normality** - Below is a histogram of the residuals. A visual inspection of the histogram indicates that the residuals are normally distributed and thus the normality assumption is met.
```{r twoWayAnova-normality}
# Normality of residuals - looks good
hist(lm3$residuals, main = "Histogram", xlab = "Residuals")
```
**Homoscedasticity** - We have used the Levene's test to check for homoscedasticity, that is to verify that the variances are equal. The null hypothesis is that the population variances are equal. The Levene's test below shows that the null hypothesis can be rejected since the p-value is very low and lower than a significance value of 0.05. Thus, there is no statistical significant evidence that the variances are equal between the experimental groups.
```{r twoWayAnova-homoscedasticity}
leveneTest(lm3)
```
### Question d.
*Present the ANOVA table. What is statistically significant? What are the effect sizes? What conclusion can you make about this experiment?*
```{r twoWayAnova-analysis, results='asis'}
lm3.anova <- anova(lm3)
lm3.anova
stargazer::stargazer(lm3.anova, header = F)
```
By analyzing the above results using the p-value, we can conclude which independent variables are statistically significant. Below is a summary of each independent variable including the interaction effect, using a significance level of 0.05.  

- Cover: p-value = 0.030075 < 0.05. It is significant.
- Summary: p-value = 0.133952 > 005. It is not significant.
- Interaction between Cover and Summary: p-value = 0.002141 < 0.05. It is significant.


**Effect Size** - The effect size, r-squared, can be calculated using:  

* SS(A) / SS(T)  
* SS(A) = factor A sum of squares  
* SS(T) = total sum of squares   

We have only calculated the effect size of the factors that were statistically significant, which was Cover and the interaction between Cover and Summary.
if we have huge data sets it's easy to find statistical significance (that's how the model works) but looking at the size of the effect tells us if this factor is relevant at all.
```{r}
TSS <- sum(lm3.anova$`Sum Sq`)
SScover <- 0.062
SScoverSummary <- 0.125
r2cover <- SScover / TSS
r2coverSummary <- SScoverSummary / TSS
c(r2cover, r2coverSummary)
```
**Conclusion** - Cover and the interaction between Cover and Summary are statistically significant which means that they impact the output value for Engagement. However, the effect size is very low for both Cover (0.15%) and the interaction (0.30%). This indicates that both factors have little to no effect on the engagement level and that other variables not included in the model influence the engagement level. This could also be seen above in the ANOVA table where the sum of squares for the residuals is much higher than both factors.

## Contributions

All four of us attended the lab. During the lab, we were able to go through the questions of the first task, discussed them and wrote down our first findings. We could not solve all aspects and only had a short look at task 2, so we split up the group: Wenli and Tim concentrated on the first task and Ludvig and Eric on the second task. After working on these parts separately, all of us checked the whole document again, so that everybody knows about all final results.
